---
title: "DATA 101 Exam 1"
author: "[ENTER NAME HERE]"
date: "Due: Monday 10/26 at 11:59pm"
output: pdf_document
---

```{r setup, include = FALSE}
# set the echo option to FALSE to see how the document looks with the code suppressed
knitr::opts_chunk$set(echo = TRUE)
```

## Academic Honesty Statement (fill in your name)

I, ____________, hereby affirm that I have not communicated with or gained information in any way from my classmates or anyone other than the Professor during this exam, that I have not assisted anyone else with this exam, and that all work is my own.

## Load packages and data

```{r load-packages, message=FALSE}
# load required packages here
```

```{r read-data, message=FALSE}
# read in the data here
```

## Questions 

### Question 1

[Enter code and narrative here.]

### Question 2

[Enter code and narrative here.]

### Question 3

[Enter code and narrative here.]

### Question 4

[Enter code and narrative here.]

### Question 5

[Enter code and narrative here.]

### Question 6

[Enter code and narrative here.]

### Question 7

[Enter code and narrative here.]

### Question 8

[Enter code and narrative here.]

### Question 9

[Enter code and narrative here.]

### Question 10

[Enter code and narrative here.]


