---
title: "DATA 101 Exam 1"
author: "Kamila Palys"
date: "Due: Monday 10/26 at 11:59pm"
output: pdf_document
---

```{r setup, include = FALSE}
# set the echo option to FALSE to see how the document looks with the code suppressed
knitr::opts_chunk$set(echo = FALSE)
```

## Academic Honesty Statement (fill in your name)

I, Kamila Palys, hereby affirm that I have not communicated with or gained information in any way from my classmates or anyone other than the Professor during this exam, that I have not assisted anyone else with this exam, and that all work is my own.

## Load packages and data

```{r load-packages, message=FALSE}
# load required packages here
library(tidyverse)
```

```{r read-data, message=FALSE}
# read in the data here
nba <- read_csv("data/nba_salaries.csv") 
```

## Questions 

### Question 1

First, we will make a table to view the salaries of the NBA players in descending order by using the "select" and "arrange" functions to display the variables we want, in the order we want.

```{r}
nba %>%
  select(player, salary) %>%
  arrange(desc(salary)) 
```

From the table it is visible that Kobe Bryant has the highest salary of all the NBA players at $25 million. 

### Question 2

Now we will create a boxplot .....
```{r}
nba 
```


### Question 3

[Enter code and narrative here.]

### Question 4

[Enter code and narrative here.]

### Question 5

[Enter code and narrative here.]

### Question 6

[Enter code and narrative here.]

### Question 7

[Enter code and narrative here.]

### Question 8

[Enter code and narrative here.]

### Question 9

[Enter code and narrative here.]

### Question 10

[Enter code and narrative here.]


