---
title: "DATA 101 Exam 1"
author: "Kamila Palys"
date: "Due: Monday 10/26 at 11:59pm"
output: pdf_document
---

```{r setup, include = FALSE}
# set the echo option to FALSE to see how the document looks with the code suppressed
knitr::opts_chunk$set(echo = FALSE)
```

## Academic Honesty Statement (fill in your name)

I, Kamila Palys, hereby affirm that I have not communicated with or gained information in any way from my classmates or anyone other than the Professor during this exam, that I have not assisted anyone else with this exam, and that all work is my own.

## Load packages and data

```{r load-packages, message=FALSE}
# load required packages here
library(tidyverse)
```

```{r read-data, message=FALSE}
# read in the data here
nba <- read_csv("data/nba_salaries.csv") 
```

## Questions 

### Question 1

First, we will make a table to view the salaries of the NBA players in descending order by using the "select" and "arrange" functions to display the variables we want, in the order we want.

```{r}
nba %>%
  select(player, salary) %>%
  arrange(desc(salary)) 
```

From the table it is visible that Kobe Bryant has the highest salary of all the NBA players at $25 million. 

### Question 2

Now we will create a boxplot that compares the distribution of the players' salaries by position. 

```{r}
ggplot(data = nba) +
  geom_boxplot(mapping = aes(x = position, y = salary)) + 
  labs(title = "NBA Salaries 2015-2016
       by Position", 
       x = "",
       y = "$ Millions per year") +
  scale_x_discrete(labels = c("Center", "Power Forward", "Point Guard", "Small Forward", "Shooting Guard"))

```

From this graph, we are able to see that the highest median salary is earned by players with the center position, while the lowest median comes from the shooting guard players. The ones who play center, however, also have a more even spread of their salaries, seeing as their interquartile range, or the middle 50%, is the largest of all the positions. There are also very little outliers in the center players' salaries, and there are several high outliers in all four of the other positions, so that is not to say that players playing a position other than center cannot earn a high salary. 

### Question 3

Now we will take a look at how many players there are in each position with "group_by" and the "count" function.

```{r}
nba %>%
  group_by(position) %>%
  count(position) %>%
  arrange(desc(position))

ggplot(data = nba, mapping = aes(x = fct_infreq(position))) +
  geom_bar() +
  labs(title = "Number of Players in each Position",
       x = "Position",
       y = "Count")
    
```

With this table and bar graph, it is visible that there are the most players in shooting guard at 96, while the center position has only 69, which is the least amount of players playing a certain position. 

### Question 4

[Enter code and narrative here.]

### Question 5

[Enter code and narrative here.]

### Question 6

[Enter code and narrative here.]

### Question 7

[Enter code and narrative here.]

### Question 8

[Enter code and narrative here.]

### Question 9

[Enter code and narrative here.]

### Question 10

[Enter code and narrative here.]


